---
title: "Initial Korean covid 19"
author: "Matthew Ramirez"
date: "Fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

When I observed the graphical plot I observed that the amount of people were alive rather than dead. However the percentage shows that more males died rather than women with women living longer than men do. The disease seems to target men by more than 2/3rds than women from less than 1/3rd.
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The Numerical results showed that the 98.7% of females were alive and only 1.4% died due to Covid while 97% of Male's survived and 3% died. Overall there were less amount of male's however, they still had a higher death rate than females.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When observing that inferential results, I found that the odds of men dying to covid in South Korea is 2.3 times higher than females, which agrees to past results that most people who were dying from COVID were men in general. I noticed that the p-value had a value of 0.0009 showing the chances of males dying from COVID rather than the females who have a lower chance from COVID.

## Conclusions

The graphical plot showed the amount of people that were apart from the statistical report and showed how more men died from Covid than women did. The numerical results confirmed that percentage amount when it showed the amount of men was almost 1/3rd more than women and how women had deaths up to 1.4% while men had up to 3% in a lower margin amount. Overall the Inferential results concluded the results with men having a ratio of 2.3 times more likely to die from Covid than women do and the p-value further concludes it with men having a 0.0009 chance of death from Covid which shows the differences in what this disease focuses on in South Korea.
